Investigating Predictors for American Attitudes on Justifiability of Abortion

### 2022-06-20 Final/FinalReport.qmd
cosmetic changes complete

---
title: "Draft"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
editor_options: 
  markdown: 
    wrap: 72
---

@@

plot_grid(h1, h2, h3, h4, h5, h6, ncol = 3, nrow = 2)

@@

plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)
```

```{r correlations}
#| warning: false
#| echo: false
ggpairs(abortion_data, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority"),
  columnLabels = c("YS", "PI", "CAI", "IG", 
                   "RA")) +
   theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 45, size = 9),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

@@

\*A note on data cleaning: Many survey questions were only added from
1995 onward, so observations before 1995 have a high number of missing
values. For these reasons, observations prior to 1995 were removed.

Year of survey, Respect for Authority, and National Pride were changed
into factors since they are categorical variables. Year has 4 levels
(one corresponding to each year), Respect for Authority has 3 levels (0
being the baseline - neutral, -1 being respect for authority bad, and 1
being respect for authority good), and National Pride is a binary (0
being the baseline - having no feelings of national pride, and 1 being
having feelings of national pride) Political Ideology, Child Autonomy
Index, and Importance of God had numerical sample spaces and were
therefore treated as numerical variables.\*

@@

```{r splits}
#| echo: false
set.seed(206)
abortion_split <- initial_split(abortion_data)
abortion_training <- training(abortion_split)
abortion_testing <- testing(abortion_split)
```

```{r spec}
#| echo: false
abortion_spec <- linear_reg() %>%
  set_engine("lm")
```

```{r recipe1}
#| echo: false
abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
                        respectauthority + nationalpride,
                        data = abortion_data) %>%
  step_center(ideology) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

@@

```{r}
df <- tidy(abortion_fit, conf.int = TRUE)
p0 <- ggplot(df, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange()
df
p0
```

@@

top_row <- plot_grid(p2,p5, ncol = 2, nrow = 1)
bottom_row <- plot_grid(p1,p3,p4, ncol = 3, nrow = 1)
plot_grid(title_factor, top_row, title_numeric, bottom_row, 
          ncol = 1, rel_heights = c(0.1, 1,0.1,1))

@@

H0 (null hypothesis): An individual's political ideology (ideology),
their beliefs about child autonomy (cai), the importance of God in their
life (godimportant), their view on whether respect for authority is good
(respectauthority), and their pride in being American (nationalpride)
are not significant predictors of aj, and the variables' respective
slope point estimates will be 0.

HA (alternative hypothesis): An individual's political ideology
(ideology), their beliefs about child autonomy (cai), the importance of
God in their life (godimportant), their view on whether respect for
authority is good (respectauthority), and their pride in being American
(nationalpride) are significant predictors of aj, and the variables'
respective slope point estimates will not be 0.

Looking at the p-values in our output displayed above, we can reject the
null hypothesis for the predictors ideology, cai, and godimportant, as
the p-values are very small, way below the 0.05 significance threshold.
Hence, we can conclude that the data provide convincing evidence that an
individual's political ideology, beliefs about child autonomy, and the
importance of God in their life are significant predictors of their
abortion attitudes.

However, we fail to reject the null hypothesis for the predictors
respectauthority and nationalpride, as the p-values for each level of
these predictors are above the 0.05 significance threshold. Hence, we
cannot conclude that the data provide convincing evidence that an
individual's view on respect for authority and their pride in being
American are significant predictors of their abortion attitudes.

@@

```{r predict-new}
#| echo: false
new_person <- tibble(year = "2011", ideology = 1, cai = 2, godimportant = 1, 
                     respectauthority = "-1", nationalpride = "0")
predict(abortion_fit, new_data = new_person)

@@

```{r metrics}
#| echo: false
abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
                                         godimportant, respectauthority))
rsq(abortion_training_pred, 
                 truth = aj, 
                 estimate = .pred)
rmse(abortion_training_pred,
                 truth = aj,
                 estimate = .pred)
```

>>>>

---
title: "Investigating Predictors for American Attitudes on Justifiability of Abortion"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
editor_options: 
  markdown: 
    wrap: 72
---

@@

title1 <- ggdraw() + 
  draw_label(
    "Distributions of Predictor variables",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

top_row1 <- plot_grid(h1,h2,h3, ncol = 3, nrow = 1)
bottom_row1 <- plot_grid(h4,h5,h6, ncol = 3, nrow = 1)
plot_grid(title1, top_row1, bottom_row1, 
          ncol = 1, rel_heights = c(0.1, 1,1))
# ADD CAPTION
```

@@

plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)

# ADD TITLE AND CAPTION
```

```{r correlations}
#| warning: false
#| echo: false
ggpairs(abortion_data, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", "nationalpride"),
  columnLabels = c("YS", "PI", "CAI", "IG", 
                   "RA", "NP")) +
   theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 45, size = 9),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
# ADD TITLE AND CAPTION
```

@@

Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values. For these reasons, observations prior to 1995 were removed.

Year of survey, Respect for Authority, and National Pride were changed
into factors since they are categorical variables. Year has 4 levels
(one corresponding to each year), Respect for Authority has 3 levels (0
being the baseline - neutral, -1 being respect for authority bad, and 1
being respect for authority good), and National Pride is a binary (0
being the baseline - having no feelings of national pride, and 1 being
having feelings of national pride) Political Ideology, Child Autonomy
Index, and Importance of God had numerical sample spaces and were
therefore treated as numerical variables.

@@

```{r splits}
#| echo: false
#| warning: false
set.seed(206)
abortion_split <- initial_split(abortion_data)
abortion_training <- training(abortion_split)
abortion_testing <- testing(abortion_split)
```

```{r spec}
#| echo: false
#| warning: false
abortion_spec <- linear_reg() %>%
  set_engine("lm")
```

```{r recipe1}
#| echo: false
#| warning: false
abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
                        respectauthority + nationalpride,
                        data = abortion_data) %>%
  step_center(ideology) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

@@

@@

top_row <- plot_grid(p2,p5, ncol = 2, nrow = 1)
bottom_row <- plot_grid(p1,p3,p4, ncol = 3, nrow = 1)
plot_grid(title_factor, top_row, title_numeric, bottom_row, 
          ncol = 1, rel_heights = c(0.1, 1,0.1,1))
#ADD CAPTION
```

@@

$H_0$ (null hypothesis): An individual's political ideology (ideology),
their beliefs about child autonomy (cai), the importance of God in their
life (godimportant), their view on whether respect for authority is good
(respectauthority), and their pride in being American (nationalpride)
are not significant predictors of aj, and the variables' respective
slope point estimates will be 0.

$H_A$ (alternative hypothesis): An individual's political ideology
(ideology), their beliefs about child autonomy (cai), the importance of
God in their life (godimportant), their view on whether respect for
authority is good (respectauthority), and their pride in being American
(nationalpride) are significant predictors of aj, and the variables'
respective slope point estimates will not be 0.

Looking at the p-values in our output displayed above, we can reject the
null hypothesis for the predictors political ideology, child autonomy index, and importance of God, as
the p-values are very small, way below the 0.05 significance threshold.
Hence, we can conclude that the data provide convincing evidence that an
individual's political ideology, beliefs about child autonomy, and the
importance of God in their life are significant predictors of their
abortion attitudes.

However, we fail to reject the null hypothesis for the predictors
respect for authority and national pride, as the p-values for each level of
these predictors are above the 0.05 significance threshold. Hence, we
cannot conclude that the data provide convincing evidence that an
individual's view on respect for authority and their pride in being
American are significant predictors of their abortion attitudes.

@@

```{r predict-new}
#| echo: false
#| eval: false
new_person <- tibble(year = "2011", ideology = 1, cai = 2, godimportant = 1, 
                     respectauthority = "-1", nationalpride = "0")
predict(abortion_fit, new_data = new_person)

@@

```{r metrics}
#| echo: false
#| eval: false
#| warning: false
abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
                                         godimportant, respectauthority))
rsq(abortion_training_pred, 
                 truth = aj, 
                 estimate = .pred)
rmse(abortion_training_pred,
                 truth = aj,
                 estimate = .pred)
```
